\begin{thebibliography}{99}

\bibitem{Alauzet-etal05}
F.~Alauzet, X.~Li, E.S. Seol, and M.S. Shephard, ``Parallel anisotropic 3d mesh adaptation by mesh modification,'' {\it Engineering with Computers,}  {\bf 21}(3):247--258 (2006).

%\bibitem{Alexandrescu01} A. Alexandrescu, Modern C++ Design: Generic Programming and Design Patterns Applied. Addison-Wesley (2001).

\bibitem{Calcote2010}
John Calcote, ``Autotools: A Practioner's Guide to GNU Autoconf, Automake, and Libtool,'' (2010).


\bibitem{Beall99} M.W. Beall, An object-oriented framework for the reliable automated solution of problems in mathematical physics [dissertation]. Troy (NY): Rensselaer Polytechnic Institute (1999).

\bibitem{Beall-Shep97}
M.W. Beall and M.S. Shephard, ``A general topology-based mesh data structure,'' {\it Int. J. Numer. Meth. Engng.,} {\bf 40}(9):1573--1596 (1997).

%\bibitem{Biswas-Oliker94}
%R. Biswas and L. Oliker, ``A new procedure for dynamic adaptation of three-dimensional
% unstructured grids,'' {\it Appl. Numer. Math.,} {\bf 13}:437--452 (1997).

\bibitem{Celes-etal05}
W. Celes, G.H. Paulino, and R. Espinha, ``A compact adjacency-based topological data structure for finite element mesh representation,'' {\it Numerical Methods in Engineering,} {\bf 64}(11):1529--1565 (2005).

%\bibitem{Cormen-etal01} T.H. Cormen, C.E. Leiserson, R.L. Rivest, and C. Stein, Introduction to Algorithms. 2nd ed. MIT Press (2001).

\bibitem{deCougny-etal95}
H.L. de Cougny, K.D. Devine, J.E. Flaherty, and R.M. Loy, ``Load balancing for the parallel solution of partial differential equations'' {\it Appl. Numer. Math.,} {\bf 16}:157--182 (1995).

\bibitem{deCougny-Shephard99}
H.L. de Cougny and M.S. Shephard, ``Parallel refinement and coarsening of tetrahedral meshes'' {\it Int. J. Numer. Meth. Engng.,} {\bf 46}:1101--1125 (1999).

%\bibitem{Deitel-Deitel01} 
%Deitel and Deitel, C++ How To Program. 2nd ed. Prentice Hall (2001).

\bibitem{Diekmann-etal00}
R.~Diekmann, R.~Preis, F.~Schlimbach, and C.~Walshaw, ``Shape-optimized mesh partitioning and load balancing for parallel adaptive fem'' {\it Parallel Computing,} {\bf 26}:1555--1581 (2000).

%\bibitem{Gamma-etal94} E. Gamma, R. Johnson, R. Helm, and J.M. Vlissides, Design Pattern: Elements of Reusable Object-Oriented Software. Addison Wesley (1994).

\bibitem{Garimella02}
R.V. Garimella, ``Mesh data structure selection for mesh generation and fea applications'' {\it Int. J. Numer. Meth. Engng.,} {\bf 55}:451--478 (2002).

%\bibitem{Ge-etal04} L. Ge, L Lee, L. Zenghai, C. Ng, K. Ko, Y Luo, and M.S. Shephard, ``Adaptive mesh refinement for high accuracy wall loss determination in accelerating cavaty design'' {\it IEEE Conf. on Electromagnetic Field Computations,} June (2004).

\bibitem{ITAPS} ITAPS: The Interoperable Technologies for Advanced Petascale Simulations center ({\it http://www.itaps.org}) (2010). 

\bibitem{fastmath} DOE SciDAC FASTMath institute ({\it http://www.fastmath-scidac.org}) (2012).

%\bibitem{Li-etal02}
%X.~Li, M.S. Shephard, and M.W. Beall, ``Accounting for curved domains in mesh adaptation'' {\it Int. J. Numer. Meth. Engng.,} {\bf 58}:247--276 (2002).

\bibitem{Li-etal03}
X.~Li, M.S. Shephard, and M.W. Beall, ``3D anisotropic mesh adaptation by mesh modifications'' {\it Comp. Meth. Appl. Mech. Engng.,} {\bf 194:}4915--4950 (2005).

\bibitem{libMesh05}
{libMesh}, {\it http://libmesh.sourceforge.net} (2005).

%\bibitem{Loy00} R. Loy, Autopack User Manual (2000).

\bibitem{Mantyla88}
M.~Mantyla, An Introduction to Solid Modeling. Rockville Maryland: Computer Science Press (1988).

\bibitem{MeshSim} Simmetrix Inc. (2005) Simulation modeling suite.
http://www.simmetrix.com.

\bibitem{Oliker-etal00}
L.~Oliker, R.~Biswas, and H.N. Gabow, ``Parallel tetrahedral mesh adaptation with dynamic load balancing'' {\it Parallel Computing,} {\bf 26}:1583--1608 (2000).

\bibitem{Ollivier-etal06}
C.~Ollivier-Gooch, K.~Chand, T.~Dahlgren, L.F.~Diachin, B.~Fix, J.~Kraftcheck, X.~Li, E.S.~Seol, M.S.~Shephard, T.~Tautges, H.~Trease ``The TSTT Mesh Interface'' {\it 44th AIAA Aerospace Sciences Meeting and Exhibit,} {\bf AIAA 2006}--529 (2006).

\bibitem{Ozturan-etal94}
C.~Ozturan, H.L. de Cougny, M.S. Shephard, and J.E. Flaherty, ``Parallel adaptive mesh refinement and redistribution on distributed memory'' {\it Comp. Meth. Appl. Mech. Engng.,} {\bf 119}:123--127 (1994).

%\bibitem{Pacheco97}
%P.S. Pacheco, Parallel Programming with MPI. Morgan Kaufmann Publisher (1997).

\bibitem{Park-Kwon05}
Y.~Park and O.~Kwon, ``A parallel unstructured dynamic mesh adaptation algorithm for 3-d
  unsteady flows'' {\it Int. J. Numer. Meth. Fluids,} {\bf 48}:671--690 (2005).

\bibitem{Remacle-etal02}
J.F. Remacle, O.~Klaas, J.E. Flaherty, and M.S. Shephard, ``A parallel algorithm oriented mesh database'' {\it Engineering with Computers,} {\bf 18}:274--284 (2002).

\bibitem{Remacle-etal03}
J.F. Remacle and M.S. Shephard, ``An algorithm oriented mesh database'' {\it Int. J. Numer. Meth. Engng.,} {\bf 58}:349--374 (2003).

%\bibitem{Scott} M.L. Scott, Programming Language Pragmatics. Kaufmann Publisher (2000).

\bibitem{Selwood-Berzins99}
P.M. Selwood and M.~Berzins, ``Parallel unstructured tetrahedral mesh adaptation: algorithms,
  implementation and scalability'' {\it Concurrency: Pract. Exper.,} {\bf 11}(14):863--884 (1999).

\bibitem{Seol05}
E.S. Seol, FMDB: Flexible Distributed Mesh Database for Parallel Automated Adaptive Analysis [dissertation]. Troy (NY): Rensselaer Polytechnic Institute (2005). Available from: {\it http://www.scorec.rpi.edu/cgi-bin/reports/GetByYear.pl?Year=2005}.

\bibitem{Seol-Shep05}
E.S. Seol and M.S. Shephard, ``Efficient distributed mesh data structure for parallel automated
  adaptive analysis'' {\it Engineering with Computers}, {\bf 22}:197--213 (2006).

\bibitem{ms:mega} 
M.S. Shephard, ``Meshing environment for
geometry-based analysis'' {\it Int. J. Numer. Meth. Engng}, 47:169-190 (2000).

\bibitem{SgiInc05} Sgi Inc. {\it http://www.sgi.com/tech/stl/stl\_index.html} (2005).

\bibitem{Shep-etal97} M.S. Shephard, J.E. Flaherty, C.L. Bottasso, and H.L. de Cougny, ``Parallel automated adaptive analysis'' {\it Parallel Computing,} {\bf 23}:1327--1347 (1997).

\bibitem{Teresco-etal00}
J.D. Teresco, M.W. Beall, J.E. Flaherty, and M.S. Shephard, ``A hierarchical partition model for adaptive finite element computations'' {\it Comp. Meth. Appl. Mech. Engng.,} {\bf 184}:269--285 (2000).

%\bibitem{Vandevoorde-Josuttis03} D. Vandevoorde and N.M. Josuttis, C++ Templates. Addison-Wesley (2003).

\bibitem{Walshaw-Cross00}
C.~Walshaw and M.~Cross, ``Parallel optimization algorithms for multilevel mesh partitioning'' {\it Parallel Computing,} {\bf 26}(12):1635--1660 (2000).

\bibitem{Weiler88}
K.J. Weiler, ``The radial-edge structure: a topological representation for
  non-manifold geometric boundary representations'' {\it Geometric Modeling for CAD Applications}, p 3--36 (1988).

\bibitem{Zoltan05} Zoltan: data management services for parallel applications (2005). Available from: {\it http://www.cs.sandia.gov/Zoltan}.

\end{thebibliography}
